---
title: "Library Mapping (nBC) Stats"
author: 'Karl Romanowicz'
output:
  html_notebook:
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_document:
    toc: yes
    toc_depth: '5'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '5'
editor_options: 
  chunk_output_type: inline
---

**GitHub Repository:** [https://github.com/PlesaLab/gene-map-nBC-docker](https://github.com/PlesaLab/gene-map-nBC-docker)

# Experiment

<font color="blue">This pipeline evaluates the mapping results from</font> <font color="red">`gene-map-nBC-docker`.</font> <font color="blue">Sequence data were generated using the Oxford Nanopore platform.</font>

```{css}
.badCode {
background-color: lightpink;
font-weight: bold;
}

.goodCode {
background-color: lightgreen;
font-weight: bold;
}

table {
  margin: auto;
  border-top: 1px solid #666;
  border-bottom: 1px solid #666;
}
table thead th { border-bottom: 1px solid #ddd; }
th, td { padding: 5px; }
thead, tfoot, tr:nth-child(even) { background: #eee; }
```

```{r setup, include=FALSE}
# Set global options for notebook
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = TRUE, message = TRUE)
knitr::opts_chunk$set(echo = TRUE, class.source = "bg-success")

# Getting the path of your current open file and set as wd
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path))
print(getwd())
```

# Packages
The following R packages must be installed prior to loading into the R session. See the **Reproducibility** tab for a complete list of packages and their versions used in this workflow.
```{r message=FALSE, results='hide', warning=FALSE}
library(dplyr)
library(readr)
library(stringi)
library(stringr)
library(tidyverse)
library('seqinr')
library(profvis)
library("stringdist")
library(gglorenz)
library(ineq)
library(patchwork)
```

```{r include=FALSE}
# set.seed is used to fix the random number generation to make the results repeatable
set.seed(123)
```

Establish a link to the SAM processing R code
```{r}
source("scripts/1_SAM_read.R")
```

# Load Data
<font color="blue">**This section is based on the R file: "Name_of_File.R".**</font> It describes how to load all of the <font color="red">pre-existing barcode mapping</font> data necessary for downstream analysis.

## Reference Data

Import REFERENCE fasta
```{r}
# 384 Reference - UNIQUE Overlaps
target_genes1 = read.fasta("refs/Unique_Twist_384-1536/unique_overlap_lib384_gene_full_wo_504.fasta",
                           seqtype = "DNA", as.string = TRUE, set.attributes=FALSE)

# 1536 Reference - UNIQUE Overlaps
# target_genes1 = read.fasta("refs/Unique_Twist_384-1536/unique_overlap_lib1536_gene_full_wo_504.fasta",
#                            seqtype = "DNA", as.string = TRUE, set.attributes=FALSE)
```

## Raw Data

Load RAW sequencing data
```{r}
# SV825S_384_unique_combined
raw1 = read.fasta("out/SV825S_384_unique_combined.fasta",
                  seqtype = "DNA", as.string = TRUE, set.attributes=FALSE)

# SV825S_1536_unique_combined
# raw1 = read.fasta("out/SV825S_1536_unique_combined.fasta",
#                   seqtype = "DNA", as.string = TRUE, set.attributes=FALSE)
```

## BBMap Files

Import SAM file from BBMap output
```{r}
# SV825S_384_unique_combined
SAM_listc1 <- read.sam("out/SV825S_384_unique_combined.bbmap.merged.sam")

# SV825S_1536_unique_combined
# SAM_listc1 <- read.sam("out/SV825S_1536_unique_combined.bbmap.merged.sam")
```

# Data Analysis

## Reference Data

```{r}
# Reformat as data frame
target_genes1 <- as.data.frame(do.call(rbind, target_genes1))
target_genes1 <- dplyr::rename(target_genes1, "dna"="V1")
target_genes1$pid <- rownames(target_genes1) #add protein IDs as column
row.names(target_genes1) <- NULL

# Capitalize DNA sequence and trim starting ATG site
target_genes1 <- target_genes1 %>%
  mutate(trimmed_dna = str_to_upper(dna))
```

```{r}
# Calculate lengths
dna_lengths <- nchar(target_genes1$dna)
trimmed_dna_lengths <- nchar(target_genes1$trimmed_dna)

# dna stats
cat("DNA Lengths:\n")
cat("Range: ", range(dna_lengths), "\n")
cat("Mean: ", mean(dna_lengths), "\n")
cat("Median: ", median(dna_lengths), "\n\n")

# trimmed_dna stats
cat("Trimmed DNA Lengths:\n")
cat("Range: ", range(trimmed_dna_lengths), "\n")
cat("Mean: ", mean(trimmed_dna_lengths), "\n")
cat("Median: ", median(trimmed_dna_lengths), "\n")
```

```{r}
# — after you’ve built target_genes1… —
target_genes1_length <- target_genes1 %>%
  mutate(ntlen = nchar(dna))
```

```{r}
# Plot histogram of ntlen
ref_seq_histogram <- ggplot(target_genes1_length, aes(x = ntlen)) +
  geom_histogram(
    binwidth = 5,
    fill     = "steelblue",
    color    = "black",
    alpha    = 0.8
  ) +
  scale_x_continuous(
    limits = c(100, 300),
    breaks = seq(100, 1050, by = 50)
  ) +
  labs(
    title = "Reference Gene Length Distribution",
    x     = "Sequence length (nt)",
    y     = "Count"
  ) +
  theme_minimal(base_size = 14)

print(ref_seq_histogram)

# save

ggsave(filename = "RMD/ref_seq_length_histogram.png", ref_seq_histogram,
  dpi = 600, width = 8, height = 6, units = "in")
```

## Raw Data

```{r}
# Reformat as data frame
raw1 <- as.data.frame(do.call(rbind, raw1))
raw1 <- dplyr::rename(raw1, "dna"="V1")
raw1$dna <- toupper(raw1$dna)
raw1$bc <- rownames(raw1) #add dna IDs as column
row.names(raw1) <- NULL

raw1 <- raw1 %>%
  mutate(ntlen=nchar(dna))
```

```{r}
# Calculate sequence lengths from the dna column
dna_lengths <- nchar(raw1$dna)

# Summary statistics for dna lengths
cat("DNA Lengths:\n")
cat("Range: ", range(dna_lengths), "\n")
cat("Mean: ", mean(dna_lengths), "\n")
cat("Median: ", median(dna_lengths), "\n\n")

# Summary statistics for ntlen
cat("ntlen Values:\n")
cat("Range: ", range(raw1$ntlen), "\n")
cat("Mean: ", mean(raw1$ntlen), "\n")
cat("Median: ", median(raw1$ntlen), "\n")
```

```{r}
# Pull out the sequence "ntlen" and "count" values for each length
raw1_gp <- raw1 %>%
  group_by(ntlen) %>%
  summarize(count=n())
```

```{r}
# Plot histogram
raw_seq_histogram <- ggplot(raw1, aes(x = ntlen)) +
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
  xlim(100,500)+
  labs(title = "Raw Sequences -- Length Distribution (504 Primer Removed)",
       x = "Sequence Length",
       y = "Frequency") +
  theme_minimal()

print(raw_seq_histogram)

ggsave(file="RMD/raw_seq_length_histogram.png", raw_seq_histogram,
       dpi=600, width = 8, height = 6, units = "in")
```

## Perfects

```{r}
perfects1 <- semi_join(raw1 %>% dplyr::rename(trimmed_dna=dna) ,target_genes1,by="trimmed_dna")

# Row Counts
row_raw <- nrow(raw1)
row_perfect <- nrow(perfects1)
perf_row_percent <- (nrow(perfects1)/nrow(raw1)) * 100

# Print results
cat("No. of Raw Data Rows:", row_raw, "\n")
cat("No. of Perfect Rows:", row_perfect, "\n")
cat("Percentage of Perfects:", round(perf_row_percent, 1), "\n")

# Basic summary statistics for the ntlen column
range_ntlen <- range(perfects1$ntlen, na.rm = TRUE)
mean_ntlen <- mean(perfects1$ntlen, na.rm = TRUE)
median_ntlen <- median(perfects1$ntlen, na.rm = TRUE)

# Print results
cat("Range of Perfects ntlen:", range_ntlen[1], "to", range_ntlen[2], "\n")
cat("Mean Perfects ntlen:", mean_ntlen, "\n")
cat("Median Perfects ntlen:", median_ntlen, "\n")
```

```{r}
#count number of perfects
target_genes1 <- perfects1 %>%
  group_by(trimmed_dna) %>%
  summarise(perfectcount1dna=n()) %>%
  right_join(target_genes1,by="trimmed_dna")
```

```{r}
#if a pid doesn't have a perfect (i.e., NA), change to 0
target_genes1$perfectcount1dna[is.na(target_genes1$perfectcount1dna)] <- 0
```

## BBMap Data

```{r}
# Convert to dataframe
aln1 <- SAM_listc1$x

# Remove SAM import to be re-used to import SAM from minimap output later
rm(SAM_listc1)
```

```{r}
# Clean up columns of interest from SAM import
aln1 <- aln1 %>%
  select(QNAME, RNAME, POS, MAPQ, CIGAR) %>%
  mutate(pid=RNAME) %>%
  filter(!str_detect(QNAME, "_part_"))

# Calculate total reads (tco1) recovered from each pid
alnsum1 <- aln1 %>%
  group_by(pid) %>%
  summarise(tco1=n())
```

```{r}
# Add tco1 counts to target_genes1 dataframe based on shared pid
target_genes1 <- target_genes1 %>%
  left_join(alnsum1, by="pid")
```

## Stats

```{r}
# Calculate percent perfect recovered for each pid
target_genes1 <- target_genes1 %>%
  mutate(pperf=perfectcount1dna/tco1*100)

# Filtered dataset retaining only pids with >99 tco1 (total reads per pid)
target_genes1_100 <- target_genes1 %>%
  filter(tco1>99)
```

```{r}
# 1. Total number of genes
total_rows <- nrow(target_genes1)
cat("Total number of genes:", total_rows, "\n")

# 2. Number of genes with at least one perfect read
nonzero_rows <- sum(target_genes1$perfectcount1dna > 0, na.rm = TRUE)
cat("Number of perfect genes (perfectcount1dna > 0):", nonzero_rows, "\n")

# 3. Percentage of genes with perfectcount1dna > 0
percent_nonzero <- (nonzero_rows / total_rows) * 100
cat("Percentage of perfect genes (perfectcount1dna > 0 / total genes):", round(percent_nonzero, 1), "\n")

# 4. Number of genes with at least 100 reads
n_100_reads <- nrow(target_genes1_100)
cat("Number of total genes with ≥100 reads:", n_100_reads, "\n")

# 5. Percentage of total gene with at least 100 reads
n_100_reads_percent <- (n_100_reads / total_rows) * 100
cat ("Percentage of total genes with > 100 reads:", round (n_100_reads_percent, 1), "\n")

# 6. Median % perfects among genes with at least 100 reads
median_pperf <- median(target_genes1_100$pperf, na.rm = TRUE)
cat("Median % perfects (genes with ≥100 reads):", round(median_pperf, 1), "\n")

# 7. Mean % perfects among genes with at least 100 reads
mean_pperf <- mean(target_genes1_100$pperf, na.rm = TRUE)
cat("Mean % perfects (genes with ≥100 reads):", round(mean_pperf, 1), "\n")

# 8. Coverage (% of genes with at least one perfect read)
coverage_percent <- (nonzero_rows / total_rows) * 100
cat("Coverage (% genes with perfectcount1dna > 0):", round(coverage_percent, 1), "%\n")

# 9. Gini coefficient for perfectcount1dna (all genes)
gini_all_perfects <- ineq(as.double(target_genes1$perfectcount1dna), type = "Gini")
cat("Gini coefficient (perfectcount1dna, all genes):", round(gini_all_perfects, 2), "\n")

# 10. Gini coefficient for perfectcount1dna (only genes with >0 perfects)
gini_nonzero_perfects <- ineq(as.double(target_genes1 %>% filter(perfectcount1dna > 0) %>% pull(perfectcount1dna)), 
                              type = "Gini")
cat("Gini coefficient (perfectcount1dna > 0):", round(gini_nonzero_perfects, 2), "\n")

# 11. Gini coefficient for total counts (tco1)
gini_total_counts <- ineq(as.double(target_genes1$tco1), type = "Gini")
cat("Gini coefficient (tco1, all genes):", round(gini_total_counts, 2), "\n")
```

```{r}
perfect_seq_histogram <- ggplot(target_genes1_100, aes(x = pperf)) +
  geom_histogram(bins = 40, color = "black", fill = "cornflowerblue") +
  theme_minimal() +
  labs(
    title = "Percentage of Perfects based on Sequence Counts",
    x = "Percentage Perfects",
    y = "Count"
  )

print(perfect_seq_histogram)

ggsave(file="RMD/perfect_histogram.png", perfect_seq_histogram,
       dpi=600, width = 8, height = 6, units = "in")
```

```{r}
# Reorder based on the perfect count of pids (high to low) - For rank plot
target_genes1 <- target_genes1 %>%
  arrange(-perfectcount1dna)

#make keys for rank plotting
target_genes1$perfects.key <- 1:length(target_genes1$perfectcount1dna)
```

```{r}
# Reorder based on the total count of pids (high to low) - For rank plot
target_genes1 <- target_genes1 %>%
  arrange(-tco1)

#make keys for rank plotting
target_genes1$tco1.key <- 1:length(target_genes1$perfectcount1dna)
```

Rank order plot based on total counts
```{r}
rank_order_total <- ggplot(target_genes1,aes(y=tco1,x=tco1.key)) +
  geom_point()+
  labs(title = "Rank Order Plot -- All Sequences",
       x = "All Reads",
       y = "Rank Ordered Gene") +
  theme(legend.position = "none")+
  scale_y_log10()+
  theme_minimal()+
  theme(legend.position = "none")

print(rank_order_total)

ggsave(file="RMD/rank_order_all.png",
       dpi=600, width = 8, height = 6, units = "in")
```

Rank order plot based on perfect counts
```{r}
rank_order_perfects <- ggplot(target_genes1,aes(y=perfectcount1dna,x=perfects.key)) +
  geom_point()+
  labs(title = "Rank Order Plot -- Perfect Sequences",
       x = "Perfect Reads",
       y = "Rank Ordered Gene") +
  theme(legend.position = "none")+
  scale_y_log10()+
  theme_minimal()+
  theme(legend.position = "none")

print(rank_order_perfects)

# Save plot
ggsave(file="RMD/rank_order_perfects.png",
       dpi=600, width = 8, height = 6, units = "in")
```

Combined Rank Order Plots
```{r}
all_plots <- (raw_seq_histogram | perfect_seq_histogram) / (rank_order_total | rank_order_perfects)
print(all_plots)

ggsave(file="RMD/All_Plots_Summary.png", all_plots,
       dpi=600, width = 12, height = 10, units = "in")
```


# Reproducibility

The session information is provided for full reproducibility.
```{r}
devtools::session_info()
```