---
title: "Library Mapping (nBC) Stats"
author: 'Karl Romanowicz'
output:
  html_notebook:
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_document:
    toc: yes
    toc_depth: '5'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '5'
editor_options: 
  chunk_output_type: inline
---

**GitHub Repository:** [https://github.com/PlesaLab/gene-map-nBC-docker](https://github.com/PlesaLab/gene-map-nBC-docker)

# Experiment

<font color="blue">This pipeline evaluates the mapping results from</font> <font color="red">`gene-map-nBC-docker`.</font> <font color="blue">Sequence data were generated using the Oxford Nanopore platform.</font>

```{r setup, include=FALSE}
# Set global options for notebook
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = TRUE, message = TRUE)
knitr::opts_chunk$set(echo = TRUE, class.source = "bg-success")

# Getting the path of your current open file and set as wd
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path))
print(getwd())
```

# Packages
The following R packages must be installed prior to loading into the R session. See the **Reproducibility** tab for a complete list of packages and their versions used in this workflow.
```{r message=FALSE, results='hide', warning=FALSE}
library(dplyr)
library(readr)
library(stringi)
library(stringr)
library(tidyverse)
library(seqinr)
library(profvis)
library(stringdist)
library(gglorenz)
library(ineq)
library(patchwork)
```

```{r include=FALSE}
# set.seed is used to fix the random number generation to make the results repeatable
set.seed(123)
```

Establish a link to the SAM processing R code
```{r}
source("scripts/1_SAM_read.R")
```

# Load Data
<font color="blue">**This section is based on the R file: "Name_of_File.R".**</font> It describes how to load all of the <font color="red">pre-existing barcode mapping</font> data necessary for downstream analysis.

## Reference Data

```{r}
# Load the REFERENCE DATA

# Catalog of reference FASTA paths (named keys)
FASTA_REF <- c(
  
    # Unique Overlaps
  Unique_lib384                    = "refs/Twist_Unique/unique_overlap_lib384_gene_full_wo_504.fasta",
  Unique_lib1536                   = "refs/Twist_Unique/unique_overlap_lib1536_gene_full_wo_504.fasta"

)

# Helper to see available keys
list_fasta_options <- function() {
  cat("Available FASTA keys:\n")
  cat(paste0(" - ", names(FASTA_REF)), sep = "\n")
}

###------------------------------------------------------------

# --- Select exactly one key to load (change this one line) ---
SELECT_FASTA <- "Unique_lib1536"

###------------------------------------------------------------

# --- Load ---
stopifnot(SELECT_FASTA %in% names(FASTA_REF))
fpath <- FASTA_REF[[SELECT_FASTA]]

if (!file.exists(fpath)) {
  stop(sprintf("File not found: %s (key: %s)", fpath, SELECT_FASTA))
}

target_genes1 <- read.fasta(
  file = fpath,
  seqtype = "DNA",
  as.string = TRUE,
  set.attributes = FALSE
)

cat(sprintf("Loaded '%s' from %s with %d sequences.\n",
            SELECT_FASTA, fpath, length(target_genes1)))
```

## Raw Data

```{r}
# Load the Raw Sequencing Data

# Catalog of raw sequencing FASTA paths (named keys)
FASTA_RAW <- c(
  
    # Unique Overlaps (Twist)---------------------------------------
  
  # Shallow Sequencing
  SV825S_384_unique_combined         = "out/SV825S_Twist_Unique_384/SV825S_Twist_Unique_384.fasta",
  SV825S_1536_unique_combined        = "out/SV825S_Twist_Unique_1536/SV825S_Twist_Unique_1536.fasta"

)

# Helper to see available keys
list_raw_options <- function() {
  cat("Available raw FASTA keys:\n")
  cat(paste0(" - ", names(FASTA_RAW)), sep = "\n")
}

###------------------------------------------------------------

# --- Select exactly one key to load (change this one line) ---
SELECT_RAW <- "SV825S_1536_unique_combined"

###------------------------------------------------------------

# --- Load ---
stopifnot(SELECT_RAW %in% names(FASTA_RAW))
raw_path <- FASTA_RAW[[SELECT_RAW]]

if (!file.exists(raw_path)) {
  stop(sprintf("File not found: %s (key: %s)", raw_path, SELECT_RAW))
}

raw1 <- read.fasta(
  file = raw_path,
  seqtype = "DNA",
  as.string = TRUE,
  set.attributes = FALSE
)

cat(sprintf("Loaded '%s' from %s with %d sequences.\n",
            SELECT_RAW, raw_path, length(raw1)))
```

## BBMap Files

```{r}
# Import SAM file from bbmap output

# --- Setup ---
#library(Rsamtools)  # or library(seqinr) if your read.sam is from seqinr

# Catalog of BBMap SAM output paths (named keys)
SAM_FILES <- c(
  
    # Unique Overlaps (Twist)---------------------------------------
  
  # Shallow Sequencing
  SV825S_384_unique_combined         = "out/SV825S_Twist_Unique_384/SV825S_Twist_Unique_384.bbmap.merged.sam",
  SV825S_1536_unique_combined        = "out/SV825S_Twist_Unique_1536/SV825S_Twist_Unique_1536.bbmap.merged.sam"

)

# Helper to see available keys
list_sam_options <- function() {
  cat("Available SAM keys:\n")
  cat(paste0(" - ", names(SAM_FILES)), sep = "\n")
}

###------------------------------------------------------------

# --- Select exactly one key to load (change this one line) ---
SELECT_SAM <- "SV825S_1536_unique_combined"

###------------------------------------------------------------

# --- Load ---
stopifnot(SELECT_SAM %in% names(SAM_FILES))
sam_path <- SAM_FILES[[SELECT_SAM]]

if (!file.exists(sam_path)) {
  stop(sprintf("File not found: %s (key: %s)", sam_path, SELECT_SAM))
}

SAM_listc1 <- read.sam(sam_path)

cat(sprintf("Loaded '%s' from %s with %d records.\n",
            SELECT_SAM, sam_path, length(SAM_listc1)))
```

# Data Analysis

## Reference Data

```{r}
# Reformat as data frame
target_genes1 <- as.data.frame(do.call(rbind, target_genes1))
target_genes1 <- dplyr::rename(target_genes1, "dna"="V1")
target_genes1$pid <- rownames(target_genes1) #add protein IDs as column
row.names(target_genes1) <- NULL

# Capitalize DNA sequence and trim starting ATG site
target_genes1 <- target_genes1 %>%
  mutate(trimmed_dna = str_to_upper(dna))
```

```{r}
# Calculate lengths
dna_lengths <- nchar(target_genes1$dna)
trimmed_dna_lengths <- nchar(target_genes1$trimmed_dna)

# dna stats
cat("DNA Lengths:\n")
cat("Range: ", range(dna_lengths), "\n")
cat("Mean: ", mean(dna_lengths), "\n")
cat("Median: ", median(dna_lengths), "\n\n")

# trimmed_dna stats
cat("Trimmed DNA Lengths:\n")
cat("Range: ", range(trimmed_dna_lengths), "\n")
cat("Mean: ", mean(trimmed_dna_lengths), "\n")
cat("Median: ", median(trimmed_dna_lengths), "\n")
```

```{r}
# — after you’ve built target_genes1… —
target_genes1_length <- target_genes1 %>%
  mutate(ntlen = nchar(dna))
```

```{r}
# Summary stats for gene counts by length
cat("\nSummary Statistics:\n")
cat("Total number of genes:", nrow(target_genes1_length), "\n")
cat("400–499 bp:", nrow(filter(target_genes1_length, ntlen >= 400 & ntlen <= 499)), "\n")
cat("500–599 bp:", nrow(filter(target_genes1_length, ntlen >= 500 & ntlen <= 599)), "\n")
cat("600–699 bp:", nrow(filter(target_genes1_length, ntlen >= 600 & ntlen <= 699)), "\n")
cat("700–799 bp:", nrow(filter(target_genes1_length, ntlen >= 700 & ntlen <= 799)), "\n")
cat("800–899 bp:", nrow(filter(target_genes1_length, ntlen >= 800 & ntlen <= 899)), "\n")
cat("900–999 bp:", nrow(filter(target_genes1_length, ntlen >= 900 & ntlen <= 999)), "\n")
cat(">=1000 bp:", nrow(filter(target_genes1_length, ntlen >= 1000)), "\n")
```

```{r}
# Plot histogram of ntlen
ref_seq_histogram <- ggplot(target_genes1_length, aes(x = ntlen)) +
  geom_histogram(
    binwidth = 5,
    fill     = "steelblue",
    color    = "black",
    alpha    = 0.8
  ) + xlim(100,1050) +
  labs(title = "Reference Gene Length Distribution",
       x = "Sequence Length",
       y = "Frequency") +
  theme_minimal()

print(ref_seq_histogram)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_ref_seq_length_histogram.png")),
  plot     = ref_seq_histogram,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

## Raw Data

```{r}
# Reformat as data frame
raw1 <- as.data.frame(do.call(rbind, raw1))
raw1 <- dplyr::rename(raw1, "dna"="V1")
raw1$dna <- toupper(raw1$dna)
raw1$bc <- rownames(raw1) #add dna IDs as column
row.names(raw1) <- NULL

raw1 <- raw1 %>%
  mutate(ntlen=nchar(dna))
```

```{r}
# Calculate sequence lengths from the dna column
dna_lengths <- nchar(raw1$dna)

# Summary statistics for dna lengths
cat("DNA Lengths:\n")
cat("Range: ", range(dna_lengths), "\n")
cat("Mean: ", mean(dna_lengths), "\n")
cat("Median: ", median(dna_lengths), "\n\n")

# Summary statistics for ntlen
cat("ntlen Values:\n")
cat("Range: ", range(raw1$ntlen), "\n")
cat("Mean: ", mean(raw1$ntlen), "\n")
cat("Median: ", median(raw1$ntlen), "\n")
```

```{r}
# Pull out the sequence "ntlen" and "count" values for each length
raw1_gp <- raw1 %>%
  group_by(ntlen) %>%
  summarize(count=n())
```

```{r}
# Plot histogram
raw_seq_histogram <- ggplot(raw1, aes(x = ntlen)) +
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
  xlim(100,1050)+
  labs(title = "Raw Sequences -- Length Distribution (504 Primer Removed)",
       x = "Sequence Length",
       y = "Frequency") +
  theme_minimal()

print(raw_seq_histogram)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_raw_seq_length_histogram.png")),
  plot     = raw_seq_histogram,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

## Perfects

```{r}
perfects1 <- semi_join(raw1 %>% dplyr::rename(trimmed_dna=dna) ,target_genes1,by="trimmed_dna")

# Row Counts
row_raw <- nrow(raw1)
row_perfect <- nrow(perfects1)
perf_row_percent <- (nrow(perfects1)/nrow(raw1)) * 100

# Print results
cat("No. of Raw Data Rows:", row_raw, "\n")
cat("No. of Perfect Rows:", row_perfect, "\n")
cat("Percentage of Perfects:", round(perf_row_percent, 1), "\n")

# Basic summary statistics for the ntlen column - Raw Reads
range_ntlen_raw <- range(raw1$ntlen, na.rm = TRUE)
mean_ntlen_raw <- mean(raw1$ntlen, na.rm = TRUE)
median_ntlen_raw <- median(raw1$ntlen, na.rm = TRUE)

# Basic summary statistics for the ntlen column - Perfect Reads
range_ntlen_perfect <- range(perfects1$ntlen, na.rm = TRUE)
mean_ntlen_perfect <- mean(perfects1$ntlen, na.rm = TRUE)
median_ntlen_perfect <- median(perfects1$ntlen, na.rm = TRUE)

cat("Range of Perfects ntlen:", range_ntlen_perfect[1], "to", range_ntlen_perfect[2], "\n")
cat("Mean Perfects ntlen:", mean_ntlen_perfect, "\n")
cat("Median Perfects ntlen:", median_ntlen_perfect, "\n")

# Print results
cat("Range of Raw ntlen:", range_ntlen_raw[1], "to", range_ntlen_raw[2], "\n")
cat("Mean Raw ntlen:", mean_ntlen_raw, "\n")
cat("Median Raw ntlen:", median_ntlen_raw, "\n")
```

```{r}
#count number of perfects
target_genes1 <- perfects1 %>%
  group_by(trimmed_dna) %>%
  summarise(perfectcount1dna=n()) %>%
  right_join(target_genes1,by="trimmed_dna")
```

```{r}
#if a pid doesn't have a perfect (i.e., NA), change to 0
target_genes1$perfectcount1dna[is.na(target_genes1$perfectcount1dna)] <- 0
```

## BBMap Data

```{r}
# Convert to dataframe
aln1 <- SAM_listc1$x

# Remove SAM import to be re-used to import SAM from minimap output later
#rm(SAM_listc1)
```

```{r}
# Clean up columns of interest from SAM import
aln1 <- aln1 %>%
  select(QNAME, RNAME, POS, MAPQ, CIGAR) %>%
  mutate(pid=RNAME) %>%
  filter(!str_detect(QNAME, "_part_"))

# Calculate total reads (tco1) recovered from each pid
alnsum1 <- aln1 %>%
  group_by(pid) %>%
  summarise(tco1=n())
```

```{r}
# Add tco1 counts to target_genes1 dataframe based on shared pid
target_genes1 <- target_genes1 %>%
  left_join(alnsum1, by="pid")
```

# Mapping Stats

```{r}
# Calculate percent perfect recovered for each pid
target_genes1 <- target_genes1 %>%
  mutate(pperf=perfectcount1dna/tco1*100)

# Filtered dataset retaining only pids with >99 tco1 (total reads per pid)
target_genes1_100 <- target_genes1 %>%
  filter(tco1>99)
```

## Library Mapping Stats

```{r}
# 1. Total number of genes
total_rows <- nrow(target_genes1)
cat("Total number of genes:", total_rows, "\n")

# 2. Number of genes with at least one perfect read
nonzero_rows <- sum(target_genes1$perfectcount1dna > 0, na.rm = TRUE)
cat("Number of perfect genes (perfectcount1dna > 0):", nonzero_rows, "\n")

# 3. Percentage of genes with perfectcount1dna > 0
percent_nonzero <- (nonzero_rows / total_rows) * 100
cat("Percentage of perfect genes (perfectcount1dna > 0 / total genes):", round(percent_nonzero, 1), "\n")

# 4. Number of total genes with at least 100 reads
n_100_reads <- nrow(target_genes1_100)
cat("Number of total genes with ≥100 reads:", n_100_reads, "\n")

# 5. Percentage of total gene with at least 100 reads
n_100_reads_percent <- (n_100_reads / total_rows) * 100
cat ("Percentage of total genes with > 100 reads:", round (n_100_reads_percent, 1), "\n")

# 6. Median % perfects among genes with at least 100 reads
median_pperf <- median(target_genes1_100$pperf, na.rm = TRUE)
cat("Median % perfects (genes with ≥100 reads):", round(median_pperf, 1), "\n")

# 7. Mean % perfects among genes with at least 100 reads
mean_pperf <- mean(target_genes1_100$pperf, na.rm = TRUE)
cat("Mean % perfects (genes with ≥100 reads):", round(mean_pperf, 1), "\n")

# 8. Coverage (% of genes with at least one perfect read)
coverage_percent <- (nonzero_rows / total_rows) * 100
cat("Coverage (% genes with perfectcount1dna > 0):", round(coverage_percent, 1), "%\n")

# 9. Gini coefficient for perfectcount1dna (all genes)
gini_all_perfects <- ineq(as.double(target_genes1$perfectcount1dna), type = "Gini")
cat("Gini coefficient (perfectcount1dna, all genes):", round(gini_all_perfects, 2), "\n")

# 10. Gini coefficient for perfectcount1dna (only genes with >0 perfects)
gini_nonzero_perfects <- ineq(as.double(target_genes1 %>% filter(perfectcount1dna > 0) %>% pull(perfectcount1dna)), type = "Gini")
cat("Gini coefficient (perfectcount1dna > 0):", round(gini_nonzero_perfects, 2), "\n")

# 11. Gini coefficient for total counts (tco1)
gini_total_counts <- ineq(as.double(target_genes1$tco1), type = "Gini")
cat("Gini coefficient (tco1, all genes):", round(gini_total_counts, 2), "\n")
```

## Global Stats

### Stats Formulas

Global Stats for Coverage and Uniformity
```{r}
gene_read_ratios <- function(gene_reads) {
  # Convert input to numeric vector
  x <- as.numeric(gene_reads)
  
  # Remove non-finite values (Inf, -Inf) and NAs
  x <- x[is.finite(x)]
  x <- x[!is.na(x)]
  
  # Count how many values remain
  n <- length(x)
  
  # Handle empty input: return NA-filled tibble if no values
  if (n == 0) return(tibble(
    ratio_90_10 = NA_real_, ratio_85_15 = NA_real_,
    frac_w_in_3x_mean  = NA_real_,
    frac_w_in_5x_mean  = NA_real_,
    frac_w_in_10x_mean = NA_real_,
    n = 0
  ))
  
  # Sort values ascending (smallest → largest) so we can slice tails
  x <- sort(x)

  # --- Tail cutpoints (indices into sorted vector) ---
  # bottom 10% index (end of lowest decile)
  i10 <- max(1L, ceiling(0.10 * n))
  # top 10% index (start of upper decile)
  i90 <- min(n, floor(0.90 * n))
  # bottom 15% index
  i15 <- max(1L, ceiling(0.15 * n))
  # top 15% index
  i85 <- min(n, floor(0.85 * n))

  # --- Mean values for tails ---
  # mean of bottom 10% of values
  mb10 <- mean(x[1:i10])
  # mean of top 10% of values
  mt10 <- mean(x[(i90 + 1):n], na.rm = TRUE)
  # mean of bottom 15% of values
  mb15 <- mean(x[1:i15])
  # mean of top 15% of values
  mt15 <- mean(x[(i85 + 1):n], na.rm = TRUE)

  # Safe division: if denominator is 0 or NA, return NA instead of error
  safe_div <- function(num, den) if (isTRUE(den > 0)) num / den else NA_real_

  # --- Tail ratios ---
  ratio_90_10 <- safe_div(mt10, mb10)  # disparity top 10% vs bottom 10%
  ratio_85_15 <- safe_div(mt15, mb15)  # disparity top 15% vs bottom 15%

  # --- Spread around mean ---
  gm <- mean(x)  # global mean
  frac_in_range <- function(k) mean(x >= gm / k & x <= gm * k)
  # fraction of values within factor of 3, 5, 10 from the mean

  # --- Return metrics as a tidy tibble ---
  tibble(
    ratio_90_10 = ratio_90_10,         # Top 10% vs bottom 10% mean ratio
    ratio_85_15 = ratio_85_15,         # Top 15% vs bottom 15% mean ratio
    frac_w_in_3x_mean  = frac_in_range(3),   # % within 3× mean
    frac_w_in_5x_mean  = frac_in_range(5),   # % within 5× mean
    frac_w_in_10x_mean = frac_in_range(10),  # % within 10× mean
    n = n                              # number of genes included
  )
}
```

### Calculated Stats

Calculate 90:10 and 85:15 ratios as well as fraction of genes within 3x, 5x, and 10x of mean:
```{r}
# After you've created target_genes1 with columns: dna, pid, perfectcount1dna, tco1

# 1. Filter down to genes that actually have at least one perfect read
# 2. Pull just the vector of perfect read counts
# 3. Pass that numeric vector into gene_read_ratios() to calculate distribution metrics
stats_perfect <- target_genes1 %>%
  dplyr::filter(perfectcount1dna > 0) %>%          # keep only genes with >0 perfect reads
  dplyr::pull(perfectcount1dna) %>%                # extract the perfectcount1dna column as a vector
  gene_read_ratios()                               # compute tail ratios + fractions near mean

# 4. Tidy up the output for readability
stats_table <- stats_perfect %>%
  mutate(
    ratio_90_10 = round(ratio_90_10, 2),           # round 90/10 ratio to 2 decimal places
    ratio_85_15 = round(ratio_85_15, 2),           # round 85/15 ratio to 2 decimal places
    frac_w_in_3x_mean  = scales::percent(          # format fraction within 3× mean as a %
      frac_w_in_3x_mean, accuracy = 0.1),
    frac_w_in_5x_mean  = scales::percent(          # format fraction within 5× mean as a %
      frac_w_in_5x_mean, accuracy = 0.1),
    frac_w_in_10x_mean = scales::percent(          # format fraction within 10× mean as a %
      frac_w_in_10x_mean, accuracy = 0.1)
  )

# 5. Print the resulting one-row tibble in the default style
#    (shows the calculated ratios, fractions, and sample size `n`)
stats_table
```

Calculate the minimum, maximum, and median lengths of perfect genes:
```{r}
dna_length_stats <- target_genes1 %>%
  # 1. Keep only genes that have at least one perfect read
  dplyr::filter(perfectcount1dna > 0) %>%
  
  # 2. Add a new column `ntlen` with the sequence length of each DNA string
  dplyr::mutate(ntlen = nchar(dna)) %>%
  
  # 3. Collapse to a one-row summary with length statistics:
  dplyr::summarise(
    min_length    = min(ntlen, na.rm = TRUE),     # shortest gene length
    max_length    = max(ntlen, na.rm = TRUE),     # longest gene length
    median_length = median(ntlen, na.rm = TRUE),  # median gene length
    span          = max_length - min_length       # difference between longest and shortest
  )

# 4. Print the summary tibble so it appears in the knitted report
dna_length_stats
```

Plot cumulative fractions within 3x, 5x, and 10x of mean as horizontal barchart:
```{r}
# numeric fractions
f3  <- stats_perfect$frac_w_in_3x_mean
f5  <- stats_perfect$frac_w_in_5x_mean
f10 <- stats_perfect$frac_w_in_10x_mean

seg_within_3x   <- f3
seg_3x_to_5x    <- pmax(f5  - f3,  0)
seg_5x_to_10x   <- pmax(f10 - f5,  0)
seg_outside_10x <- pmax(1   - f10, 0)

# desired conceptual order
order_levels <- c("Within 3× mean", "Additional 3–5×", "Additional 5–10×", "Outside 10×")

plot_df <- tibble::tibble(
  band = factor(
    c("Within 3× mean", "Additional 3–5×", "Additional 5–10×", "Outside 10×"),
    levels = order_levels
  ),
  frac = c(seg_within_3x, seg_3x_to_5x, seg_5x_to_10x, seg_outside_10x)
) |>
  dplyr::mutate(
    label = scales::percent(frac, accuracy = 0.1),
    show  = frac >= 0.02
  )

# Auto-detect dataset name from fasta_file path
dataset_name <- tools::file_path_sans_ext(basename(SELECT_RAW))

#file.path("RMD", paste0(SELECT_RAW, "_raw_seq_length_histogram.png")),

# Make the stacked bar chart
p_dist_stack <- ggplot(plot_df, aes(x = "Coverage bands", y = frac, fill = band)) +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  geom_text(
    data = dplyr::filter(plot_df, show),
    aes(label = label),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    color = "white", size = 3.6, fontface = "bold"
  ) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1), expand = c(0, 0)) +
  scale_fill_discrete(limits = order_levels, guide = guide_legend(reverse = FALSE)) +
  labs(
    title = paste("Distribution of genes around the mean coverage —", dataset_name),
    x = NULL, y = "Fraction of genes", fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

p_dist_stack

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_gene_distribution_mean_coverage.png")),
  plot     = p_dist_stack,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

# Library Plots

## Length Histogram

```{r}
perfect_seq_histogram <- ggplot(target_genes1_100, aes(x = pperf)) +
  geom_histogram(bins = 40, color = "black", fill = "cornflowerblue") +
  theme_minimal() +
  labs(
    title = "Percentage of Perfects based on Sequence Counts",
    x = "Percentage Perfects",
    y = "Count"
  )

print(perfect_seq_histogram)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_perfect_histogram.png")),
  plot     = perfect_seq_histogram,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

## Rank Order Plots

```{r}
# Reorder based on the perfect count of pids (high to low) - For rank plot
target_genes1 <- target_genes1 %>%
  arrange(-perfectcount1dna)

#make keys for rank plotting
target_genes1$perfects.key <- 1:length(target_genes1$perfectcount1dna)
```

```{r}
# Reorder based on the total count of pids (high to low) - For rank plot
target_genes1 <- target_genes1 %>%
  arrange(-tco1)

#make keys for rank plotting
target_genes1$tco1.key <- 1:length(target_genes1$perfectcount1dna)
```

Rank order plot based on total counts
```{r}
rank_order_total <- ggplot(target_genes1,aes(y=tco1,x=tco1.key)) +
  geom_point()+
  labs(title = "Rank Order Plot -- All Sequences",
       x = "All Reads",
       y = "Rank Ordered Gene") +
  theme(legend.position = "none")+
  scale_y_log10()+
  theme_minimal()+
  theme(legend.position = "none")

print(rank_order_total)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_rank_order_all.png")),
  plot     = rank_order_total,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

Rank order plot based on perfect counts
```{r}
rank_order_perfects <- ggplot(target_genes1,aes(y=perfectcount1dna,x=perfects.key)) +
  geom_point()+
  labs(title = "Rank Order Plot -- Perfect Sequences",
       x = "Perfect Reads",
       y = "Rank Ordered Gene") +
  theme(legend.position = "none")+
  scale_y_log10()+
  theme_minimal()+
  theme(legend.position = "none")

print(rank_order_perfects)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_rank_order_perfects.png")),
  plot     = rank_order_perfects,
  width    = 8,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)
```

## Combined Plots

```{r}
all_plots <- (raw_seq_histogram | perfect_seq_histogram) / (rank_order_total | rank_order_perfects)
print(all_plots)

# Save plot with sample key in the filename
ggsave(
  filename = file.path("RMD", paste0(SELECT_RAW, "_all_plots_summary.png")),
  plot     = all_plots,
  width    = 12,
  height   = 10,
  dpi      = 300,
  bg       = "white"
)
```

# Reproducibility

The session information is provided for full reproducibility.
```{r}
devtools::session_info()
```